---
title: "Life Expectancy for Political Candidates"
author: "Afnan K"
format: html
execute: 
  echo: false
---

```{r}
#| label: fit-years-model
#| cache: true

library(tidymodels)
library(modeldata)

data(governors)
x <- governors

fit_years <- linear_reg(engine = "lm") |>
  fit(lived_after ~ sex + election_age, data = x)
```


```{r setup}
#| message: false
library(tidyverse)
library(primer.data)
library(tidymodels)
library(broom)
library(gt)
library(marginaleffects)

x <- governors

```

```{r}
# Load required libraries
library(tidyverse)

# Set up a tibble of ages
ages <- seq(35, 80, by = 1)

# Coefficients from your model
intercept <- 19.32338863
age_coef <- -0.05631269
sexMale_coef <- 53.29554022
interaction_coef <- -0.79997911

# Predictions
predictions <- tibble(
  election_age = ages,
  Female = intercept + age_coef * election_age,
  Male = intercept + age_coef * election_age + sexMale_coef + interaction_coef * election_age
) |>
  pivot_longer(cols = c(Female, Male), names_to = "sex", values_to = "predicted_years")

# Plot
ggplot(predictions, aes(x = election_age, y = predicted_years, color = sex)) +
  geom_line(linewidth = 1.2) +
  labs(
    title = "Predicted Years Lived After Election by Sex and Age at Election",
    x = "Age at Election",
    y = "Predicted Years Lived After Election",
    color = "Sex"
  ) +
  theme_minimal(base_size = 14)


```


Candidates' lives are often affected by elections. Using data from the paper "Longevity Returns to Political Office" by Barfort, Klemmensen & Larsen (2019), which includes all candidates in U.S. gubernatorial elections after 1945, we seek to create a model for forecasting longevity for political candidates. Because our data includes only candidates for governor, our estimated model might not generalize well to candidates for other political offices, such as mayor or city council. We model life expectancy after the election, measured in years lived after election day, as a linear function of a candidate’s sex and age at the time of election. A 50-year-old male political candidate can expect to live about 30 years after the election, plus-or-minus about 1 year.




## Model Structure

We represent our statistical model as:

$$
Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_k X_k + \varepsilon
$$


## Fitted Model

The estimated linear regression model is:

$$
\hat{lived\_after} = 58.3 + 0.15 \cdot sex_{Male} - 0.68 \cdot election\_age
$$


## Regression Model Estimates

```{r}
#| label: regression-table
#| cache: true

# Extract model estimates as a tibble with confidence intervals
fit_summary <- tidy(fit_years, conf.int = TRUE)

# Create a formatted regression table
fit_summary |>
  select(term, estimate, conf.low, conf.high) |>
  gt() |>
  tab_header(
    title = "Regression Model Estimates",
    subtitle = "Predicting Years Lived After Election"
  ) |>
  fmt_number(
    columns = c(estimate, conf.low, conf.high),
    decimals = 2
  ) |>
  cols_label(
    term = "Term",
    estimate = "Estimate",
    conf.low = "Lower 95% CI",
    conf.high = "Upper 95% CI"
  ) |>
  tab_source_note(
    source_note = "Source: 'governors' dataset from Barfort et al. (2019)"
  ) |>
  tab_caption(
    caption = "Table: Estimated coefficients and 95% confidence intervals from a linear regression predicting years lived after election."
  )
```





